Loop Invariant Code Motion in PyPy's Tracing JIT

### 2011-06-16 talk/iwtc11/paper.tex
we have no guard_true any more

Loop peeling is achieved by appending an inlined copy of the traced iteration at
the end of itselfe. See Figure~\ref{fig:overview}.
The first part (called \emph{preamble}) finishes with the jump the the second part
(called the \emph{peeled loop}). The second part end with the jump to itself. This way
the preamble will be executed only once while the peeled loop will
be used for every further iteration.

When applying the following optimizations to this two-iteration trace
some care has to taken as to how the arguments of the two
\lstinline{jump} operations and the input arguments of the peeled loop are
treated. It has to be ensured that the peeled loop stays a proper
trace in the sense that the operations within it only operates on
variables that are either among its input arguments 
or produced within the peeled loop. To ensure this we need
to introduce a bit of formalism.

The original trace (prior to peeling) consists of three parts.
A vector of input
variables, $I=\left(I_1, I_2, \cdots, I_{|I|}\right)$, a list of non-
jump operations and a single
jump operation. The jump operation contains a vector of jump variables,
$J=\left(J_1, J_2, \cdots, J_{|J|}\right)$, that are passed as the input variables of the target loop. After
loop peeling there will be a second copy of this trace with input
variables equal to the jump arguments of the preamble, $J$, and jump
arguments $K$. Looking  at the peeled version of our example in Figure~\ref{fig:peeled-trace} we have
\begin{equation}
  %\left\{
    \begin{array}{lcl}
      I &=& \left( p_0, p_1 \right) \\
      J &=& \left( p_0, p_5 \right) \\
      K &=& \left( p_0, p_9 \right) \\
    \end{array}
  %\right.
  .
\end{equation}
To construct the second iteration from the first we also need a
function $m$, mapping the variables of the preamble onto the
variables of the peeled loop. This function is constructed during the
inlining. It is initialized by mapping the input arguments, $I$, to
the jump arguments $J$,
\begin{equation}
  m\left(I_i\right) = J_i \ \text{for}\ i = 1, 2, \cdots |I| .
\end{equation}
In the example that means:

>>>>

Loop peeling is achieved by appending an copy of the traced iteration at
the end of itselfe. See Figure~\ref{fig:overview}.
The first part (called \emph{preamble}) finishes with the jump the the second part
(called the \emph{peeled loop}). The second part end with the jump to itself. This way
the preamble will be executed only once while the peeled loop will
be used for every further iteration. New variable names have to be
introduced in the entire copied trace in order to maintian the SSA-property.
When applying the following optimizations to this two-iteration trace
some care has to taken as to how the arguments of the two
\lstinline{jump} operations and the input arguments of the peeled loop are
treated. It has to be ensured that the peeled loop stays a proper
trace in the sense that the operations within it only operates on
variables that are either among its input arguments 
or produced within the peeled loop. To ensure this we need
to introduce a bit of formalism.

The original trace (prior to peeling) consists of three parts.
A vector of input
variables, $I=\left(I_1, I_2, \cdots, I_{|I|}\right)$, a list of non-
jump operations and a single
jump operation. The jump operation contains a vector of jump variables,
$J=\left(J_1, J_2, \cdots, J_{|J|}\right)$, that are passed as the input variables of the target loop. After
loop peeling there will be a second copy of this trace with input
variables equal to the jump arguments of the preamble, $J$, and jump
arguments $K$. Looking  at the peeled version of our example in Figure~\ref{fig:peeled-trace} we have
\begin{equation}
  %\left\{
    \begin{array}{lcl}
      I &=& \left( p_0, p_1 \right) \\
      J &=& \left( p_0, p_5 \right) \\
      K &=& \left( p_0, p_9 \right) \\
    \end{array}
  %\right.
  .
\end{equation}
To construct the second copy of the trace (the peeled loop) from the
first (the preeamble) we need a
function $m$, mapping the variables of the preamble onto the
variables of the peeled loop. This function is constructed during the
copying. It is initialized by mapping the input arguments, $I$, to
the jump arguments $J$,
\begin{equation}
  m\left(I_i\right) = J_i \ \text{for}\ i = 1, 2, \cdots |I| .
\end{equation}
In the example that means:
